---
title: "Freight Emissions Monte Carlo Summary"
format: html
---

```{r}
#| echo: false
options(stringsAsFactors = FALSE)
```

## Problem Statement

This report summarizes a distributed Monte Carlo simulation of freight emissions for dry vs refrigerated distribution.
The system boundary is **Manufacturing → Retail Transportation** and the functional unit is **1,000 kcal delivered to retail**.

## Methods

Deterministic emissions per functional unit:

```
Emissions_gCO2 = (Mass_per_FU_kg / 907.185) * Distance_miles *
  (Truck_EF + Reefer_Penalty) * Utilization_Factor
```

Uncertainty is propagated via Monte Carlo sampling, and aggregation uses exact moment merging plus mergeable histograms.

## Assumptions

```{r}
assumptions_path <- "data/inputs_local/assumptions_used.csv"
assumptions <- if (file.exists(assumptions_path)) read.csv(assumptions_path) else data.frame()
assumptions
```

## Results

```{r}
agg_path <- "outputs/aggregate/results_summary.csv"
local_path <- "outputs/local/results_summary.csv"

results_path <- if (file.exists(agg_path)) agg_path else local_path
results <- if (file.exists(results_path)) read.csv(results_path) else data.frame()
results
```

```{r}
# Plot histogram if aggregate hist exists
hist_path <- "outputs/aggregate/hist_diff_gco2.csv"
if (file.exists(hist_path)) {
  h <- read.csv(hist_path)
  mids <- (h$bin_left + h$bin_right) / 2
  plot(mids, h$count, type = "h", main = "diff_gco2 Histogram", xlab = "gCO2", ylab = "Count")
}
```
