---
title: "Transportation and Cold-Chain Implications: Proposal-Aligned Results"
format: html
---

```{r}
#| echo: false
options(stringsAsFactors = FALSE)
```

## Research Questions

1. How much additional GHG is introduced by refrigerated vs dry freight, normalized to **1,000 kcal delivered to retail**?
2. How do outcomes vary across **CENTRALIZED vs REGIONALIZED** spatial structures?
3. How do outcomes vary across **Diesel tractor + diesel TRU** vs **BEV tractor + electric TRU load**?
4. What is `P(refrigerated > dry)` for each scenario/powertrain combination?
5. At what distance scale does refrigeration become emissions-dominant?

## Model Framing

The simulation runs in an offline-first Monte Carlo pipeline with deterministic seeds.  
Functional unit is first-class via `data/inputs_local/functional_unit.csv` (`FU_1000_KCAL`).

Per draw, outputs include:

- `ghg_traction`
- `ghg_refrigeration` (zero for dry mode)
- `ghg_total = ghg_traction + ghg_refrigeration`

## Inputs and Assumptions

```{r}
assumptions_path <- "data/inputs_local/assumptions_used.csv"
assumptions <- if (file.exists(assumptions_path)) read.csv(assumptions_path) else data.frame()
assumptions
```

```{r}
fu_path <- "data/inputs_local/functional_unit.csv"
fu <- if (file.exists(fu_path)) read.csv(fu_path) else data.frame()
fu
```

## Scenario Results (Dry vs Refrigerated)

```{r}
cmp_path <- "outputs/analysis/scenario_comparison.csv"
cmp <- if (file.exists(cmp_path)) read.csv(cmp_path, stringsAsFactors = FALSE) else data.frame()
if (nrow(cmp) == 0) {
  cat("No proposal scenario comparison found. Run:\n")
  cat("`Rscript tools/summarize_proposal_outputs.R --runs_dir outputs/proposal --outdir outputs/analysis`\n")
} else {
  cmp
}
```

## Probability Refrigerated Exceeds Dry

```{r}
if (nrow(cmp) > 0) {
  old_par <- par(mar = c(8, 4, 2, 1))
  on.exit(par(old_par), add = TRUE)
  labels <- paste(cmp$spatial_structure, cmp$powertrain_config, sep = " | ")
  barplot(
    cmp$p_refrigerated_gt_dry,
    names.arg = labels,
    las = 2,
    col = "#3D7EA6",
    ylim = c(0, 1),
    ylab = "Probability"
  )
  abline(h = 0.5, lty = 2, col = "gray40")
}
```

## Distance Threshold Sensitivity

```{r}
th_path <- "outputs/analysis/distance_thresholds.csv"
ths <- if (file.exists(th_path)) read.csv(th_path, stringsAsFactors = FALSE) else data.frame()
if (nrow(ths) == 0) {
  cat("No threshold output found yet.\n")
} else {
  ths
}
```

## Route Simulation (Offline Planned Stops)

```{r}
rs_path <- "outputs/analysis/route_sim_summary_stats.csv"
rs <- if (file.exists(rs_path)) read.csv(rs_path, stringsAsFactors = FALSE) else data.frame()
if (nrow(rs) == 0) {
  cat("No route simulation summary found. Run: `make route-sim-summary`\\n")
} else {
  rs
}
```

## Traffic Emissions Penalty (Paired)

```{r}
tep_pair_path <- "outputs/analysis/route_sim_traffic_penalty.csv"
tep_sum_path <- "outputs/analysis/route_sim_traffic_penalty_summary.csv"
tep_pair <- if (file.exists(tep_pair_path)) read.csv(tep_pair_path, stringsAsFactors = FALSE) else data.frame()
tep_sum <- if (file.exists(tep_sum_path)) read.csv(tep_sum_path, stringsAsFactors = FALSE) else data.frame()
if (nrow(tep_sum) == 0) {
  cat("No paired TEP summary found yet.\\n")
} else {
  tep_sum
}
```

## Geography Sensitivity Index (Protein-Normalized, Paired)

```{r}
gsi_pair_path <- "outputs/analysis/route_sim_geo_sensitivity_protein.csv"
gsi_sum_path <- "outputs/analysis/route_sim_geo_sensitivity_protein_summary.csv"
gsi_pair <- if (file.exists(gsi_pair_path)) read.csv(gsi_pair_path, stringsAsFactors = FALSE) else data.frame()
gsi_sum <- if (file.exists(gsi_sum_path)) read.csv(gsi_sum_path, stringsAsFactors = FALSE) else data.frame()
if (nrow(gsi_sum) == 0) {
  cat("No paired GSI summary found yet.\\n")
} else {
  gsi_sum
}
```

```{r}
sens_path <- "outputs/analysis/distance_sensitivity.csv"
sens <- if (file.exists(sens_path)) read.csv(sens_path, stringsAsFactors = FALSE) else data.frame()
if (nrow(sens) > 0) {
  groups <- unique(paste(sens$scenario_id, sens$powertrain_config, sep = "|"))
  cols <- c("#4C78A8", "#F58518", "#54A24B", "#E45756")
  plot(NA, NA,
    xlim = range(sens$distance_multiplier, na.rm = TRUE),
    ylim = range(sens$delta_p50, na.rm = TRUE),
    xlab = "Distance Multiplier",
    ylab = "Median Delta (Refrigerated - Dry)"
  )
  abline(h = 0, lty = 2, col = "gray40")
  for (i in seq_along(groups)) {
    g <- groups[[i]]
    ss <- sens[paste(sens$scenario_id, sens$powertrain_config, sep = "|") == g, , drop = FALSE]
    lines(ss$distance_multiplier, ss$delta_p50, col = cols[((i - 1) %% length(cols)) + 1], lwd = 2)
  }
  legend("topleft", legend = groups, col = cols[seq_along(groups)], lwd = 2, bty = "n")
}
```
