---
title: "Cold-Chain Freight Emissions Monte Carlo Simulation Report"
author: "Generated by coldchainfreight"
date: "`r Sys.Date()`"
format:
  html:
    toc: true
    toc-depth: 3
    code-fold: true
    theme: cosmo
    embed-resources: true
  pdf:
    toc: true
    number-sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(ggplot2)
library(coldchainfreight)
```

## Executive Summary

This report presents results from a distributed Monte Carlo simulation comparing emissions from dry versus refrigerated freight transport.

```{r load-results}
# Load simulation results
summary_stats <- readRDS("outputs/summary_stats.rds")
histogram_dry <- readRDS("outputs/histogram_dry.rds")
histogram_refrigerated <- readRDS("outputs/histogram_refrigerated.rds")
merged_moments <- readRDS("outputs/merged_moments.rds")
artifact <- jsonlite::read_json("outputs/contribution_artifact.json")
```

### Key Findings

- **Total Samples Simulated**: `r format(summary_stats$total_samples, big.mark = ",")`
- **Dry Freight Samples**: `r format(summary_stats$n_dry, big.mark = ",")`
- **Refrigerated Freight Samples**: `r format(summary_stats$n_refrigerated, big.mark = ",")`

## Emission Statistics

### Overall CO₂ Emissions

- **Mean**: `r round(summary_stats$overall_mean_co2, 2)` kg
- **Standard Deviation**: `r round(summary_stats$overall_sd_co2, 2)` kg
- **Variance**: `r round(summary_stats$overall_variance, 2)`

### Dry Freight

- **Mean CO₂**: `r round(summary_stats$dry_mean_co2, 2)` kg
- **Std Dev**: `r round(summary_stats$dry_sd_co2, 2)` kg
- **Mean Cost**: $`r round(summary_stats$dry_mean_cost, 2)`

### Refrigerated Freight

- **Mean CO₂**: `r round(summary_stats$refrig_mean_co2, 2)` kg
- **Std Dev**: `r round(summary_stats$refrig_sd_co2, 2)` kg
- **Mean Cost**: $`r round(summary_stats$refrig_mean_cost, 2)`

### Cost Impact

The additional cost of refrigeration averages **$`r round(summary_stats$cost_difference, 2)`** per shipment.

## Emission Distribution Comparison

```{r emission-distributions, fig.width=10, fig.height=6}
# Create data frame for plotting
plot_data <- data.frame(
  emissions = c(
    rep(histogram_dry$mids, histogram_dry$counts),
    rep(histogram_refrigerated$mids, histogram_refrigerated$counts)
  ),
  type = c(
    rep("Dry Freight", sum(histogram_dry$counts)),
    rep("Refrigerated", sum(histogram_refrigerated$counts))
  )
)

# Plot distributions
ggplot(plot_data, aes(x = emissions, fill = type)) +
  geom_histogram(alpha = 0.6, position = "identity", bins = 40) +
  scale_fill_manual(values = c("Dry Freight" = "#2E86AB", "Refrigerated" = "#A23B72")) +
  labs(
    title = "CO₂ Emissions Distribution: Dry vs Refrigerated Freight",
    x = "CO₂ Emissions (kg)",
    y = "Frequency",
    fill = "Freight Type"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
    legend.position = "top"
  )
```

```{r emission-density, fig.width=10, fig.height=6}
# Density plot
ggplot(plot_data, aes(x = emissions, color = type, fill = type)) +
  geom_density(alpha = 0.3, size = 1.2) +
  scale_color_manual(values = c("Dry Freight" = "#2E86AB", "Refrigerated" = "#A23B72")) +
  scale_fill_manual(values = c("Dry Freight" = "#2E86AB", "Refrigerated" = "#A23B72")) +
  labs(
    title = "CO₂ Emissions Density: Dry vs Refrigerated Freight",
    x = "CO₂ Emissions (kg)",
    y = "Density",
    color = "Freight Type",
    fill = "Freight Type"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
    legend.position = "top"
  )
```

## Statistical Moments

The following table shows exact statistical moments calculated using the parallel axis theorem for accurate aggregation across distributed chunks:

```{r moments-table}
moments_df <- data.frame(
  Statistic = c("Sample Size", "Mean", "Variance", "Standard Deviation", "Skewness", "Kurtosis"),
  Value = c(
    format(merged_moments$n, big.mark = ","),
    round(merged_moments$mean, 4),
    round(merged_moments$variance, 4),
    round(merged_moments$sd, 4),
    round(merged_moments$skewness, 4),
    round(merged_moments$kurtosis, 4)
  )
)

knitr::kable(moments_df, caption = "Exact Statistical Moments")
```

## Simulation Parameters

```{r params-table}
params_df <- data.frame(
  Parameter = c(
    "Distance Mean (km)",
    "Distance SD (km)",
    "Payload Mean (tons)",
    "Payload SD (tons)",
    "Temperature Mean (°C)",
    "Temperature SD (°C)",
    "Fuel Efficiency Mean (L/100km)",
    "Fuel Efficiency SD (L/100km)",
    "Refrigeration Probability"
  ),
  Value = c(
    artifact$parameters$distance_mean,
    artifact$parameters$distance_sd,
    artifact$parameters$payload_mean,
    artifact$parameters$payload_sd,
    artifact$parameters$temp_mean,
    artifact$parameters$temp_sd,
    artifact$parameters$fuel_efficiency_mean,
    artifact$parameters$fuel_efficiency_sd,
    artifact$parameters$refrigeration_prob
  )
)

knitr::kable(params_df, caption = "Simulation Input Parameters")
```

## Reproducibility Information

- **Simulation ID**: `r artifact$simulation_id`
- **Base Random Seed**: `r artifact$base_seed`
- **Number of Chunks**: `r artifact$n_chunks`
- **Chunk Size**: `r format(artifact$chunk_size, big.mark = ",")`
- **Reproducibility Hash**: `r artifact$reproducibility_hash`
- **Timestamp**: `r artifact$timestamp`

This simulation was designed for full reproducibility. All random number generation uses explicit seeds, and the reproducibility log captures all execution details. To reproduce these results, use the same base seed and parameters with the `targets` pipeline.

## Methodology

### Deterministic Emission Model

The emission model calculates fuel consumption and CO₂ emissions based on:

1. **Base fuel consumption**: Distance × fuel efficiency
2. **Payload adjustment**: +2% fuel consumption per ton
3. **Refrigeration overhead**: 25% base increase + temperature-dependent adjustment
4. **CO₂ conversion**: 2.68 kg CO₂ per liter of diesel

### Monte Carlo Sampling

- Chunks of samples are generated with explicit random seeds
- Input parameters follow normal distributions
- Each chunk is independently executable for distributed computing

### Aggregation Methods

- **Histograms**: Counts are summed across chunks with consistent breaks
- **Moments**: Exact merging using parallel axis theorem
- **Results**: Combined data frames from all chunks

## Conclusion

This Monte Carlo simulation provides a rigorous comparison of emissions between dry and refrigerated freight transport. The results demonstrate the additional environmental and cost burden of cold-chain logistics, while maintaining strict reproducibility standards suitable for research applications.

---

*Report generated by coldchainfreight package*
