---
title: "Getting Started with coldchainfreight"
author: "coldchainfreight package"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with coldchainfreight}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5
)
```

## Introduction

The `coldchainfreight` package implements a research-grade Monte Carlo simulation for comparing emissions from dry versus refrigerated (cold-chain) freight transport. This vignette demonstrates the basic workflow.

## Installation

```{r eval=FALSE}
# Install from GitHub
devtools::install_github("dMac716/coldchain-freight-montecarlo")
```

## Basic Emission Calculations

The core function is `calculate_emissions()`, which computes fuel consumption, CO₂ emissions, and costs:

```{r eval=FALSE}
library(coldchainfreight)

# Dry freight
dry_result <- calculate_emissions(
  distance_km = 500,
  payload_tons = 20,
  is_refrigerated = FALSE,
  ambient_temp_c = 20,
  fuel_efficiency_l_per_100km = 30
)

print(dry_result)
```

### Understanding the Model

The emission model accounts for:

1. **Base fuel consumption**: Calculated from distance and fuel efficiency
2. **Payload adjustment**: Additional 2% fuel per ton
3. **Refrigeration overhead**: 25% base increase plus temperature effects
4. **CO₂ conversion**: 2.68 kg CO₂ per liter of diesel

## Monte Carlo Simulation

### Single Chunk

A "chunk" is an independent batch of Monte Carlo samples:

```{r eval=FALSE}
# Define parameter distributions
params <- list(
  distance_mean = 500, distance_sd = 100,
  payload_mean = 20, payload_sd = 5,
  temp_mean = 20, temp_sd = 8,
  fuel_efficiency_mean = 30, fuel_efficiency_sd = 5,
  refrigeration_prob = 0.3
)

# Initialize reproducibility log
init_reproducibility_log("my_simulation.json")

# Run one chunk
chunk <- run_mc_chunk(
  chunk_id = 1,
  chunk_size = 1000,
  seed = 12345,
  params = params
)

# View results
summary(chunk$results$co2_emissions_kg)
```

### Multiple Chunks

For larger simulations, run multiple chunks and merge:

```{r eval=FALSE}
# Run 3 chunks
chunks <- lapply(1:3, function(i) {
  run_mc_chunk(
    chunk_id = i,
    chunk_size = 1000,
    seed = 12345 + i * 1000,
    params = params
  )
})

# Merge statistical moments
moments_list <- lapply(chunks, function(c) c$moments)
merged_moments <- merge_moments(moments_list)

print(merged_moments)
```

## Exact Statistical Merging

The package uses exact formulas for combining statistics:

### Moment Merging

Uses the parallel axis theorem:

```{r eval=FALSE}
# Moments from two chunks merge exactly
merged <- merge_moments(list(chunk1$moments, chunk2$moments))

# These are exact, not approximate
merged$mean      # Exact combined mean
merged$variance  # Exact combined variance
merged$skewness  # Combined skewness
```

### Histogram Aggregation

Histograms with consistent breaks can be merged:

```{r eval=FALSE}
# Aggregate histograms by freight type
hist_dry <- aggregate_histograms(chunks, "dry")
hist_refrigerated <- aggregate_histograms(chunks, "refrigerated")

# Plot
plot(hist_dry$mids, hist_dry$density, type = "l",
     main = "Dry Freight CO₂ Distribution",
     xlab = "CO₂ (kg)", ylab = "Density")
```

## Using the Targets Pipeline

For production simulations, use the included `targets` pipeline:

```{r eval=FALSE}
library(targets)

# Run the pipeline
tar_make()

# View results
tar_read(summary_stats)
tar_read(merged_moments)

# Render report
tar_make(report)
```

The pipeline handles:
- Chunk generation with seeds
- Parallel execution (if configured)
- Result merging
- Report generation
- Artifact saving with JSON schema validation

## Reproducibility

### Logging

All operations are logged for reproducibility:

```{r eval=FALSE}
# Initialize log
init_reproducibility_log("simulation.json")

# Log is automatically updated
run_mc_chunk(1, 1000, 12345, params)

# Check reproducibility hash
hash <- get_reproducibility_hash()
```

### Validation

The package enforces:
- All random seeds are explicit
- All parameters are validated
- No runtime API calls (offline operation)
- Complete session information is logged

## Input Validation

All functions validate inputs:

```{r eval=FALSE}
# This will error with clear message
calculate_emissions(
  distance_km = -100,  # Invalid
  payload_tons = 20,
  is_refrigerated = FALSE,
  ambient_temp_c = 20,
  fuel_efficiency_l_per_100km = 30
)
# Error: Parameter 'distance_km' must be >= 0 (got -100)
```

## Best Practices

### For Research Use

1. **Always initialize logging**: `init_reproducibility_log()`
2. **Use explicit seeds**: Document all random seeds used
3. **Save contribution artifacts**: Use the JSON schema
4. **Document parameters**: Record all simulation parameters
5. **Version control**: Use git to track parameter changes

### For Large Simulations

1. **Choose appropriate chunk size**: 1,000-10,000 samples
2. **Use distributed computing**: via `targets` with `clustermq` or `future`
3. **Save only aggregated results**: Don't save all raw samples
4. **Use histogram binning**: More memory efficient than raw data

### For Collaboration

1. **Share seeds and parameters**: Others can reproduce exactly
2. **Use contribution artifacts**: Validated JSON format
3. **Include reproducibility hash**: Verify execution environment
4. **Document deviations**: Log any parameter changes

## Advanced Topics

### Custom Parameter Distributions

Modify `run_mc_chunk()` to use different distributions:

```{r eval=FALSE}
# The function uses rnorm() by default
# You can modify to use other distributions:
# - rlnorm() for log-normal
# - rgamma() for gamma
# - runif() for uniform
```

### Parallel Execution with Future

```{r eval=FALSE}
library(future)
library(future.apply)

# Set up parallel backend
plan(multisession, workers = 4)

# Run chunks in parallel
results <- future_lapply(1:10, function(i) {
  run_mc_chunk(i, 1000, 12345 + i * 1000, params)
})

plan(sequential)  # Reset
```

### Sensitivity Analysis

```{r eval=FALSE}
# Vary a parameter and observe effects
temps <- seq(10, 40, by = 5)

sensitivity_results <- lapply(temps, function(temp) {
  p <- params
  p$temp_mean <- temp
  
  chunk <- run_mc_chunk(1, 500, 12345, p)
  
  list(
    temp = temp,
    mean_co2 = chunk$moments$mean,
    sd_co2 = sqrt(chunk$moments$variance)
  )
})

# Plot sensitivity
df <- do.call(rbind, lapply(sensitivity_results, as.data.frame))
plot(df$temp, df$mean_co2, type = "l",
     main = "Temperature Sensitivity",
     xlab = "Mean Temperature (°C)",
     ylab = "Mean CO₂ (kg)")
```

## Conclusion

The `coldchainfreight` package provides a robust framework for Monte Carlo simulation of freight emissions with:

- Deterministic, validated models
- Exact statistical aggregation
- Full reproducibility
- Distributed computing support
- Research-grade implementation

For more information, see the package documentation and examples.
