---
title: "Route Map: Factory to Retail"
---

```{r, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}
if (!requireNamespace("leaflet", quietly = TRUE)) {
  cat("<p><em>Leaflet package not available. Map rendering skipped.</em></p>")
  knitr::knit_exit()
}
suppressPackageStartupMessages({
  library(leaflet)
  library(htmlwidgets)
  library(htmltools)
})

# Null-coalesce operator (needed for fac_colors lookup)
`%||%` <- function(a, b) if (!is.null(a) && !is.na(a) && nzchar(a)) a else b

pick_path <- function(paths) {
  for (p in paths) if (file.exists(p)) return(p)
  ""
}

# Source only the polyline decoder (sourcing all R/ files errors on missing deps)
geo_path <- pick_path(c("../../R/route_geometry_utils.R", "../R/route_geometry_utils.R"))
if (nzchar(geo_path)) {
  tryCatch(source(geo_path, local = TRUE), error = function(e) {
    cat(sprintf("<p><em>Failed to load route geometry utils: %s</em></p>", e$message))
  })
}

routes_path <- pick_path(c(
  "../../data/derived/routes_facility_to_petco.csv",
  "../data/derived/routes_facility_to_petco.csv",
  "data/derived/routes_facility_to_petco.csv"
))

facilities_path <- pick_path(c(
  "../../data/inputs_local/facilities.csv",
  "../data/inputs_local/facilities.csv"
))

retail_path <- pick_path(c(
  "../../data/inputs_local/retail_nodes.csv",
  "../data/inputs_local/retail_nodes.csv"
))

if (!nzchar(routes_path)) {
  cat("<p><em>Route data not found. Run <code>bash tools/route_precompute_google.sh</code> to generate routes.</em></p>")
  knitr::knit_exit()
}

routes <- utils::read.csv(routes_path, stringsAsFactors = FALSE)

# Facility metadata
fac <- if (nzchar(facilities_path)) utils::read.csv(facilities_path, stringsAsFactors = FALSE) else data.frame()
ret <- if (nzchar(retail_path)) utils::read.csv(retail_path, stringsAsFactors = FALSE) else data.frame()

# Decode polylines into lat/lon paths
decode_route <- function(encoded_polyline) {
  tryCatch(decode_polyline(encoded_polyline), error = function(e) NULL)
}

# Color scheme: Topeka (dry) = blue, Ennis (refrigerated) = orange
fac_colors <- c(
  FACILITY_DRY_TOPEKA = "#0d4f8f",
  FACILITY_REFRIG_ENNIS = "#e97a28"
)
fac_labels <- c(
  FACILITY_DRY_TOPEKA = "Topeka, KS (Dry)",
  FACILITY_REFRIG_ENNIS = "Ennis, TX (Refrigerated)"
)

# Only show rank-1 routes (primary) with full opacity; alternatives lighter
map <- leaflet(width = "100%", height = "620px") |>
  addProviderTiles("OpenStreetMap.Mapnik", group = "Street") |>
  addProviderTiles("CartoDB.Positron", group = "Light") |>
  addProviderTiles("Esri.NatGeoWorldMap", group = "Terrain") |>
  addLayersControl(
    baseGroups = c("Street", "Light", "Terrain"),
    options = layersControlOptions(collapsed = FALSE)
  )

for (i in seq_len(nrow(routes))) {
  r <- routes[i, , drop = FALSE]
  pts <- decode_route(as.character(r$encoded_polyline))
  if (is.null(pts) || nrow(pts) < 2) next

  fid <- as.character(r$facility_id)
  col <- fac_colors[fid] %||% "#666666"
  label <- fac_labels[fid] %||% fid
  is_primary <- as.integer(r$route_rank) == 1L
  dist_mi <- round(as.numeric(r$distance_m) / 1609.344)
  dur_h <- round(as.numeric(r$duration_s) / 3600, 1)
  dur_static_h <- round(as.numeric(r$duration_s_static) / 3600, 1)
  pref <- as.character(r$routing_preference)

  popup_txt <- sprintf(
    paste0(
      "<b>%s</b><br/>",
      "<b>Route %d</b> to Petco Davis<br/>",
      "Distance: %s mi<br/>",
      "Duration: %s h (traffic) / %s h (static)<br/>",
      "Routing: %s"
    ),
    label, r$route_rank, format(dist_mi, big.mark = ","),
    dur_h, dur_static_h, pref
  )

  map <- map |>
    addPolylines(
      lng = pts$lon, lat = pts$lat,
      color = col,
      weight = if (is_primary) 4 else 2,
      opacity = if (is_primary) 0.85 else 0.35,
      dashArray = if (is_primary) NULL else "8,6",
      popup = popup_txt,
      group = label
    )
}

# Permanent label styling
perm_label_style <- labelOptions(
  noHide = TRUE, direction = "top", textOnly = FALSE,
  style = list("font-weight" = "bold", "font-size" = "11px",
               "background" = "rgba(255,255,255,0.85)", "padding" = "2px 6px",
               "border-radius" = "3px")
)

# Short labels for permanent display
fac_short_labels <- c(
  FACILITY_DRY_TOPEKA = "Topeka (Dry)",
  FACILITY_REFRIG_ENNIS = "Ennis (Refrig)"
)

# Add facility markers with permanent labels
if (nrow(fac) > 0) {
  for (j in seq_len(nrow(fac))) {
    f <- fac[j, , drop = FALSE]
    fid <- as.character(f$facility_id)
    col <- fac_colors[fid] %||% "#333"
    lab <- fac_labels[fid] %||% fid
    short_lab <- fac_short_labels[fid] %||% fid
    map <- map |>
      addCircleMarkers(
        lng = as.numeric(f$lon), lat = as.numeric(f$lat),
        radius = 10, color = col, fillColor = col, fillOpacity = 0.9,
        stroke = TRUE, weight = 2,
        label = htmltools::HTML(short_lab),
        labelOptions = perm_label_style,
        popup = sprintf("<b>%s</b><br/>%s, %s<br/>%.4f, %.4f",
                        lab, f$city, f$state, f$lat, f$lon)
      )
  }
}

# Add retail marker with permanent label
if (nrow(ret) > 0) {
  r <- ret[1, , drop = FALSE]
  map <- map |>
    addCircleMarkers(
      lng = as.numeric(r$lon), lat = as.numeric(r$lat),
      radius = 10, color = "#d62728", fillColor = "#d62728", fillOpacity = 0.9,
      stroke = TRUE, weight = 2,
      label = htmltools::HTML("Petco Davis"),
      labelOptions = perm_label_style,
      popup = sprintf("<b>%s</b><br/>%s, %s<br/>Retail destination",
                      r$retail_id, r$city, r$state)
    )
}

# Add legend
map <- map |>
  addLegend(
    "bottomright",
    colors = c(fac_colors, "#d62728"),
    labels = c(fac_labels, "Petco Davis (Retail)"),
    title = "Locations",
    opacity = 1
  )

map
```

::: {.panel}
### Route Details

This map displays the actual Google Routes API driving routes from each distribution facility to the retail destination (Petco, Davis CA). Routes are computed with **TRAFFIC_AWARE_OPTIMAL** routing preference, providing both traffic-adjusted and static (free-flow) durations.

- **Blue routes**: Topeka, KS (dry product facility) --- 1,712 miles
- **Orange routes**: Ennis, TX (refrigerated product facility) --- 1,774 miles
- **Solid lines**: Primary (rank 1) route; **dashed**: alternative routes
- Click any route segment for distance, duration, and routing metadata

These routes are the geometric basis for the BEV charging waypoint plans and the Monte Carlo transport simulation.
:::

::: {.panel}
### Route Usage in Production Audit (59,434 runs)

```{r, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}
audit_stats_path <- pick_path(c(
  "../assets/transport/audit_2026-03-17/tables/comprehensive_scenario_stats.csv",
  "../../site/assets/transport/audit_2026-03-17/tables/comprehensive_scenario_stats.csv"
))
if (nzchar(audit_stats_path)) {
  astats <- utils::read.csv(audit_stats_path, stringsAsFactors = FALSE)
  # Aggregate runs by origin_network
  origin_agg <- aggregate(n_runs ~ origin_network, data = astats, FUN = sum)
  origin_agg <- origin_agg[order(-origin_agg$n_runs), ]

  cat("| Origin Network | Total Runs | Powertrains | Mean Distance (mi) | Mean CO2 (kg) |\n")
  cat("|:---|---:|:---|---:|---:|\n")
  for (k in seq_len(nrow(origin_agg))) {
    on <- origin_agg$origin_network[k]
    subset_rows <- astats[astats$origin_network == on, ]
    pts <- paste(sort(unique(subset_rows$powertrain)), collapse = ", ")
    avg_dist <- round(weighted.mean(subset_rows$mean_distance_miles, subset_rows$n_runs), 1)
    avg_co2 <- round(weighted.mean(subset_rows$mean_co2_kg_total, subset_rows$n_runs), 1)
    cat(sprintf("| %s | %s | %s | %s | %s |\n",
                on, format(origin_agg$n_runs[k], big.mark = ","), pts,
                format(avg_dist, big.mark = ","), format(avg_co2, big.mark = ",")))
  }
  cat(sprintf("\n**Total runs**: %s across %d scenario combinations.\n",
              format(sum(astats$n_runs), big.mark = ","),
              nrow(astats)))
} else {
  cat("Audit data not yet available. Place `comprehensive_scenario_stats.csv` in `assets/transport/audit_2026-03-17/tables/`.\n")
}
```
:::

::: {.panel}
### FAF Freight Flow Context

The simulation also draws from USDA Freight Analysis Framework (FAF5) origin-destination flow data to model realistic distance distributions for the centralized and regionalized spatial scenarios.

```{r, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}
flows_path <- pick_path(c(
  "../../data/derived/faf_top_od_flows.csv",
  "../data/derived/faf_top_od_flows.csv"
))
if (nzchar(flows_path)) {
  flows <- utils::read.csv(flows_path, stringsAsFactors = FALSE)
  cat(sprintf("- **%d** origin-destination pairs from FAF5\n", nrow(flows)))
  cat(sprintf("- Scenarios: %s\n", paste(unique(flows$scenario_id), collapse = ", ")))
  cat(sprintf("- Distance range: %.0f -- %.0f miles\n", min(flows$distance_miles, na.rm = TRUE), max(flows$distance_miles, na.rm = TRUE)))
} else {
  cat("FAF flow data not available.\n")
}
```
:::
